**---
title: "Social Science"
output:
  pdf_document: default
  html_document: html_document
date: "2023-05-01"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


# QUESTION 1 A
# READ CSV FILE
```{r }

df <- read.csv2("D:\\Social ScienceProject #3 (1)\\Project\\shiw_2016.csv")
str(df)

```
```{r }
summary(df)

```

## GROUPING ETA
```{r }
df$age_class <- cut(df$eta, breaks = c(0, 30, 65, Inf), labels = c("young", "adult", "old"))
str(df)
```

## QUESTION 2 B
# GROUPING CONSUMTION
```{r }
min(df$c)
max(df$c)

# finding the mid consumption
threshold <- min(df$c[df$c > min(df$c)])
threshold

df$T_c <- cut(df$c, breaks=c(min(df$c), threshold, max(df$c)), labels=c("Low","High"))
df$T_c
```
### QUESTION 2 C
# COMPUTING CONTINGENCY TABLE
```{r }
conti <- table(df$age_class, df$T_c)
conti 
```

## MARGINAL DISTRIBUTION ACCORDING TO age_class
```{r }
marginal <- margin.table(conti,margin=1)
marginal
```

### JOINT DISTRIBUTION ACCORDING TO age_class

```{r }
joint <- prop.table(conti,margin=1)
joint
```
#### conditional distribution of T_c given age_class

```{r }
condit <- prop.table(conti,margin=2)
condit
```
##### finding high consumption by older people

```{r }
older_high_consumption <- conti["old", "High"]
older_high_consumption
```
.
*The number of people that showed high level of consumption is 2960 according to the consumption level threshold made above*
.

#### QUESTION 2 C 
# HYPOTHESIS
```{r }

chi <- chisq.test(table(df$age_class, df$T_c))
chi
```
```{r }
if (chi$p.value < 0.05) {
  print("Reject null hypothesis: age_class and T_c are associated")
} else {
  print("Do not reject the null hypothesis: age_class and T_c are independent")
}
```
.
*we do not reject the null hypothesis since the p-value is greater than the 5% significance threshold, therefore 
there is no enough evidence to prove that age_class and T_c are associated*
.