**---
title: "Analysis on FActors Affecting Global Crude Oil"
output:
  pdf_document: default
  html_document: default
date: "2023-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Preprocessing
```{r }
da <- read.csv("First.csv")
df<-da
str(df)
```


# Correlation
```{r }
cor(df)
library(ggcorrplot)
ggcorrplot(cor(df))

library(psych)
library(ggplot2)
pairs.panels(df)
```


# Scatter Plot
```{r }
library(ggplot2)
dd<- ggplot(data=df, aes(x=Oprod, y=Oprice ))
dd +geom_point()+ggtitle("Distribution of Oil price and Oil Production")

dd2<- ggplot(data=df, aes(x=Gprice, y=Oprice))
dd2 +geom_point() +ggtitle("Distribution of Oil price and Gas Price")

dd3<- ggplot(data=df, aes(x=Gprod, y=Oprice))
dd3 +geom_point() + ggtitle("Distribution of Oil price and Gas Production")

dd4<- ggplot(data=df, aes(x=Oreserve, y=Oprice))
dd4 +geom_point() +ggtitle("Distribution of Oil price and Oil Proved Reserve")

dd5<- ggplot(data=df, aes(x=Wrate, y=Oprice))
dd5 +geom_point() + ggtitle("Distribution of Oil price and World Economy")

```




# Multiple Regression

```{r }
set.seed(100)
```

## First model

```{r }
mo <- lm(Oprice~., data=df)
summary(mo)
mean(mo$residuals^2)
```




### Multicolinearity
```{r }
library(GGally)
ggpairs(df)
```

# 2nd model
```{r  }
model2 <- lm(Oprice~ Gprice+Oprod+Wrate+Oreserve, data=df)
summary(model2)
mean(model2$residuals^2)

```

# 3rd model
## Backward AIC
```{r }
model3=step(mo,direction = "backward") 
mean(model3$residuals^2)
```

```{r }
summary(model3)


```



# Forward AIC
```{r }
model4=step(mo, direction="forward")

mean(model4$residuals^2)

```
```{r }
summary(model4)
```

# BOth FOrward and BAckward
```{r }
model5=step(mo, direction="both")
summary(model5)
mean(model5$residuals^2)

```
## Mean square error
```{r }
mean(model2$residuals^2)
mean(model3$residuals^2)
mean(model4$residuals^2)
mean(model5$residuals^2)
```


# Anova
```{r }
anova(mo, model2 )

```
# Bagging 

```{r }

set.seed(1)
idx_tr = sample(nrow(df), 0.8*nrow(df))
df_train = df[idx_tr, ]
df_test = df[-idx_tr,]

```

```{r }
library(randomForest)
bag = randomForest(Oprice~.,data=df_train, mtry=5, importance=TRUE )
bag
names(bag)
summary(bag)
```

## Calculatingn the importance
```{r }
importance(bag)
varImpPlot(bag)

```

### Prediction
```{r }
Predict_bag = predict(bag, df_test)
plot(Predict_bag, df_test$Oprice)
abline(0,1)
```


#### MEan Square error
```{r }
MSE_bag= mean((Predict_bag-df_test$Oprice)^2)
MSE_bag

```


```{r }
bag2 = randomForest(Oprice~.,data=df_train, mtry=5, ntree=50, importance=TRUE )
bag2
summary(bag2)
Predict_bag2 = predict(bag2, df_test)
plot(Predict_bag2, df_test$Oprice)
abline(0,1)
MSE_bag2= mean((Predict_bag2-df_test$Oprice)^2)
MSE_bag2
```

# randomForest
```{r }
MSE.Rf=rep(0,5)
for(d in 1:5){
  rf.df=randomForest(Oprice~., data=df_train, mtry=d, importance=TRUE)

  predict.rf=predict(rf.df, df_test)

  MSE.Rf[d]=mean((predict.rf-df_test$Oprice)^2)

}
MTRY=c(1:5)


```
## Plotting MSE for each size of ntree 

```{r }
plot(MTRY, MSE.Rf, type="b", col="red")
min(MSE.Rf)
data.frame(MTRY,MSE.Rf)
```

### randomForest
```{r }
set.seed(400)
rf.data=randomForest(Oprice~., data=df_train, mtry=2, importance=TRUE)
rf.data
summary(rf.data)
```

#### Prediction RF of mtry=2
```{r }
predict_rf.data=predict(rf.data, df_test)
MSE_rf_mtry2=mean((predict_rf.data-df_test$Oprice))
MSE_rf_mtry2
```

##### Importance
```{r }
importance(rf.data)
varImpPlot(rf.data)
```




# MTRY=3
```{r }
set.seed(400)
rf.data2=randomForest(Oprice~., data=df_train, mtry=3, importance=TRUE)
rf.data2
summary(rf.data2)
```

#### Prediction RF of mtry=3
```{r }
predict_rf.data2=predict(rf.data2, df_test)
MSE_rf_mtry3=mean((predict_rf.data2-df_test$Oprice))
MSE_rf_mtry3
```

##### Importance
```{r }
importance(rf.data2)
varImpPlot(rf.data2)
```

```{r }

importance <- data.frame(Importance = rf.data2$importance[order(rf.data2$importance[,1], decreasing = TRUE), 1])
rownames(importance) <- rownames(rf.data2$importance)[order(rf.data2$importance[,1], decreasing = TRUE)]
ggplot(importance, aes(x = rownames(importance), y = Importance)) + 
  geom_bar(stat="identity", fill="steelblue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
**