---
title: "Linear Regression"
author: "Your Name"
date: "2023-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(MASS)
data("Boston", package="MASS")
str(Boston)
```
# Numeric to factor
```{r}
Boston$chas=factor(Boston$chas, labels=c("No", "Yes"))
summary(Boston)
```

#Predictong the median house value
```{r}
Boston$medv
require(tidyverse)

Boston %>%
    select(!chas) %>%
    cor()

jpeg("D:\\MACHINE LEARNING\\Anyinam\\th\\medv_cor.jpeg")
{
Boston %>%
    select(!chas) %>%
    pairs()    
}
dev.off()
```

.
*Medv is highly correlated with rm(average number of rooms) and lstat(percentage of households with low socioeconomic status*
.

```{r}
require(gridExtra)

jpeg("D:\\MACHINE LEARNING\\Anyinam\\th\\scatter_plot_rm_lstat.jpeg")
p1 = ggplot(Boston, aes(x=lstat, y=medv))+ 
geom_point(alpha=0.5, col="blue")

p2=ggplot(Boston, aes(x=rm, y=medv))+
geom_point(alpha=0.5, col="green")

grid.arrange(p1,p2)
dev.off()

```

## Simple Linear Regression
```{r}
m1 =lm(medv~lstat, data=Boston)
summary(m1)

```

```{r}
jpeg("D:\\MACHINE LEARNING\\Anyinam\\th\\medv_abline.jpeg")
par(mfrow=c(1,1))

plot(Boston$lstat,Boston$medv,pch=19,cex=0.5, xlab="lstat", ylab="medv") 
abline(m1,col="red")
dev.off()
```
```{r}

coefficients(m1)
```
.
*The increasing of a one unit of the percentage of poor households in an area 
is associated with a decreasing of approximately 950$ of the median house values (on average)*
.

```{r}
confint(m1, level = 0.95)

```